---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Google-COVID-19-Mobility-Reports

<!-- badges: start -->
<!-- badges: end -->

In this notebook I use `R` to retrieve Google's [community mobility reports](https://www.google.com/covid19/mobility/). These reports are being released by Google to support public health experts and the public understand the way mobility has changed with respect to a baseline level for the duration of the COVID-19 pandemic. I was inspired to do this after seeing Theo Lim's GitHub [repository](https://github.com/theochli/google_covid_mobility_map) with Python code to download the reports and visualize the data. This is mostly an example of how to retrieve data, and I will add analysis as I can.

Reports are presented by location with variations; for instance, Canada's community mobility reports are given at the national and provincial levels. For some countries the reports are given only at the national level. For the US the reports are at the level of counties. Each report includes the percent change in visits to places like grocery stores and parks within the designated geographic area. The baseline is the median value, for the corresponding day of the week, during the 5-week period between January, 3rd and February 6th, 2020. Given seasonality in mobility, it would probably have been better to use a baseline of the corresponding day of the week for a moving window of 5 or whatever weeks, but of the previous year. I wonder if those data from previous years even exist.

The categories of places are as follows:

- Grocery & pharmacy: Mobility trends for places like grocery markets, food warehouses, farmers markets, specialty food shops, drug stores, and pharmacies.

- Parks: Mobility trends for places like local parks, national parks, public beaches, marinas, dog parks, plazas, and public gardens.

- Transit stations: Mobility trends for places like public transport hubs such as subway, bus, and train stations.

- Retail & recreation: Mobility trends for places like restaurants, cafes, shopping centers, theme parks, museums, libraries, and movie theaters.

- Residential: Mobility trends for places of residence.

- Workplaces: Mobility trends for places of work.

More information about the data can be found [here](https://www.google.com/covid19/mobility/data_documentation.html?hl=en).

## Preliminaries

Load the packages needed to read and work with the data:
```{r message=FALSE}
library(countrycode)
library(gridExtra)
library(interactions)
library(jtools)
library(lubridate)
library (RCurl)
library(tidyverse)
library(wppExplorer)
```

Retrieve data:
```{r}
download <- getURL("https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv?cachebust=911a386b6c9c230f")
gmobility <- read.csv (text = download)
```

Save as RData. I created a folder called `data` to keep a local copy of the mobility report:
```{r}
save(gmobility, file = "data/gmobility.RData")
```

What are the contents of the report?
```{r}
head(gmobility)
```

The unit of the report is a combination of region and date, and the columns include the percent change from baseline for each category of place.

Check the range of dates covered:
```{r}
summary(gmobility$date)
```

The earliest date reported for any geographical location February 15, 2020. However, the `date` is stored as a factor which is not ideal work with times; I will change it to date:
```{r}
gmobility <- gmobility %>%
  mutate(date = lubridate::ymd(date))
```

Summarize the dates again:
```{r}
summary(gmobility$date)
```

That is better. Now it is clear that the earliest date is Feb. 15, 2020 (the minimum value) and the latest (for the time being) is `r paste(month(max(gmobility$date), label= TRUE), ". ", day(max(gmobility$date)), ", ", year(max(gmobility$date)), sep = "")`.

With this more or less the community reports are in workable shape.

## How has mobility in Canada changed according to Google?

To begin an exploration of mobility in Canada (according to Google), the first step is to filter the dataframe to include only Canadian data (I'm calling this "cmr" for "community mobility report"):
```{r}
canada_cmr <- gmobility %>%
  filter(country_region == "Canada") %>%
  droplevels()
```

We can quickly check the contents of this Canadian dataframe:
```{r}
summary(canada_cmr)
```

Now we have an All-Canadian dataset with the highest geographical level of resolution the province. At this point, I am going to rename the annoyingly long names of the columns to something more practical (after all, I know that all these variables are "percent change from baseline"):
```{r}
canada_cmr <- canada_cmr %>%
  rename(r_and_r = retail_and_recreation_percent_change_from_baseline, 
         g_and_p = grocery_and_pharmacy_percent_change_from_baseline, 
         parks = parks_percent_change_from_baseline,
         transit = transit_stations_percent_change_from_baseline, 
         work = workplaces_percent_change_from_baseline,
         residential = residential_percent_change_from_baseline)
```

Another thing that we notice is that there are quite a few "NAs" in the table. These could be data that have been censored by Google to comply with their [Privacy Policy](https://www.google.com/covid19/mobility/index.html?hl=en). Maybe. But that seems like a little bit of a stretch, given that the aggregation is at the provincial level. Or it could be that in fact there are relatively few people in some provinces at certain times who allow Google to track them. Or could the "NAs" be that on certain dates there were absolutely no trips to those places? What would be the percentage change from the baseline to zero?

We can explore this question by looking at where the "NAs" are:
```{r}
canada_cmr %>% 
  filter(is.na(r_and_r)) %>%
  select(sub_region_1) %>%
  table()
```

Ah. So the "NAs" are concentrated in the Northwest Territories, Nunavut, and Yukon, all sparsely populated regions; likely the "NAs" then are a result of censoring due to small numbers. Notice that there is also an unlabeled "region". This are in fact the national-level statistics.

We can verify that this is the case too with the other categories of places, like groceries and pharmacies:
```{r}
canada_cmr %>% 
  filter(is.na(g_and_p)) %>%
  select(sub_region_1) %>%
  table()
```

Parks:
```{r}
canada_cmr %>% 
  filter(is.na(parks)) %>%
  select(sub_region_1) %>%
  table()
```

In addition to the three regions identified before, Prince Edward Island also has "NAs" in the category of parks.

Transit:
```{r}
canada_cmr %>% 
  filter(is.na(transit)) %>%
  select(sub_region_1) %>%
  table()
```

Work:
```{r}
canada_cmr %>% 
  filter(is.na(work)) %>%
  select(sub_region_1) %>%
  table()
```

Residential:
```{r}
canada_cmr %>% 
  filter(is.na(work)) %>%
  select(sub_region_1) %>%
  table()
```

Now, there are different ways in which these data could be analyzed (obviously), but I will begin with data at the national level. I could remove the provinces/territories with "NAs", or I could include them and ignore the records that are missing. I'll do the latter.

Create a plot of the changes in mobility with respect to retail and recreation, taking the daily average of all provinces (taking care of removing the "NAs" from the calculation of the mean, so using na.rm = TRUE):
```{r}
canada_cmr %>%
  group_by(date) %>%
  summarize (r_and_r = mean(r_and_r, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = r_and_r)) +
  geom_line() +
  theme_bw() +
  ylab("% change wrt baseline: retail and recreation")
```

We see that, at the national level, the mobility hovered close to the baseline (percent change ~ 0) until March 13, which is when social distancing measures started to be phased in across the country. Mobility with respect to retail and recreation after that date collapsed to about -50% change with respect to the baseline, and hovered there (with additional drops in the second week of April).

But did this change similarly across the country? Repeat the plot, but by macro-region. I will add the macro-regions to the dataframe:
```{r}
canada_cmr <- canada_cmr %>%
  mutate(macroregion = case_when(sub_region_1 == "Ontario" ~ "Ontario",
                                 sub_region_1 == "Quebec" ~ "Quebec",
                                 sub_region_1 == "British Columbia" ~ "British Columbia",
                                 sub_region_1 == "Manitoba" | 
                                   sub_region_1 == "Saskatchewan" | 
                                   sub_region_1 == "Alberta"  ~ "Prairies",
                                 sub_region_1 == "New Brunswick" | 
                                   sub_region_1 == "Newfoundland and Labrador" | 
                                   sub_region_1 == "Nova Scotia" | 
                                   sub_region_1 == "Prince Edward Island"  ~ "Atlantic",
                                 sub_region_1 == "Northwest Territories" | 
                                   sub_region_1 == "Nunavut" | 
                                   sub_region_1 == "Yukon" ~ "Northern Canada"),
         macroregion = factor(macroregion))
```

```{r}
canada_cmr %>%
  drop_na(macroregion) %>%
  group_by(date, macroregion) %>%
  summarize (r_and_r = mean(r_and_r, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = r_and_r, color = macroregion)) +
  geom_line() +
  theme_bw() +
  ylab("% change wrt baseline: retail and recreation")
```

Here, we see that Quebec has had the largest changes in retail-and-recreation-related mobility of all regions, whereas Northern Canada has had the least (albeit still quite considerable)(but we don't know what was the baseline; if the base line was low mobility, then there was maybe not much that could be cut).

With most form of mobility severely curtailed by social distancing mandates, residential-based mobility is perhaps the form of mobility that dominates during the pandemic. The following plot shows how residential-based mobility has changed (as before, this is the daily average of all provinces, after removing the "NAs" from the calculation of the mean):
```{r}
canada_cmr %>%
  group_by(date) %>%
  summarize (residential = mean(residential, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = residential)) +
  geom_line() +
  theme_bw() +
  ylab("% change wrt baseline: retail and recreation")
```

Whereas retail and recreation-related mobility has collapsed, residential-based mobility as increased following the orders that restricted non-essential travel. This is not necessarily surprising: time-use research by Canadian researchers [Harvey and Taylor (2000)](https://link.springer.com/article/10.1023/A:1005207320044) had shown that working from home did not necessarily reduce travel, but instead shifted its purpose. These researchers argue that people tend to seek social contact, and when social contact is not available through attendance to a workplace, place of study, etc., that they will shift their travel to other matters.

What where the changes in residential-based mobility by region? 
```{r}
canada_cmr %>%
  drop_na(macroregion) %>%
  group_by(date, macroregion) %>%
  summarize (residential = mean(residential, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = residential, color = macroregion)) +
  geom_line() +
  theme_bw() +
  ylab("% change wrt baseline: residential")
```

Interestingly, we see that residential-based mobility grew more in the place where retail and recreation mobility dropped most.

## COVID-19 cases from Johns Hopkins University

```{r}
baseURL = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series"

covid19 = read.csv(file.path("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series",
                           "time_series_covid19_confirmed_global.csv"),
                 check.names=FALSE, 
                 stringsAsFactors=FALSE) %>%
  select(-Lat, -Long) %>% 
  pivot_longer(-(1:2), names_to = "date", values_to = "Cases") %>% 
  mutate(date=as.Date(date, format="%m/%d/%y"),
         `Country/Region`=if_else(`Country/Region` == "", "?", `Country/Region`),
         `Province/State`=if_else(`Province/State` == "", "<all>", `Province/State`))
```

Save as RData. This is going to the same folder where I saved the community mobility reports:
```{r}
save(covid19, file = "data/covid19_jhu.RData")
```

Filter data for Canada and convert variables to factors/date, as appropriate:
```{r}
Country <- "Canada"
country_code <- countrycode(Country, origin = "country.name", destination = "iso2c")
canada_c19 <- covid19 %>%
  filter(`Country/Region` == Country) %>%
  transmute(Province = factor(`Province/State`),
            Country = factor(`Country/Region`),
            date = ymd(date),
            Cases)
```

We can also add to the dataframe the macro-regions:
```{r}
canada_c19 <- canada_c19 %>%
  mutate(macroregion = case_when(Province == "Ontario" ~ "Ontario",
                                 Province == "Quebec" ~ "Quebec",
                                 Province == "British Columbia" ~ "British Columbia",
                                 Province == "Manitoba" | 
                                   Province == "Saskatchewan" | 
                                   Province == "Alberta"  ~ "Prairies",
                                 Province == "New Brunswick" | 
                                   Province == "Newfoundland and Labrador" | 
                                   Province == "Nova Scotia" | 
                                   Province == "Prince Edward Island"  ~ "Atlantic",
                                 Province == "Northwest Territories" | 
                                   Province == "Nunavut" | 
                                   Province == "Yukon" ~ "Northern Canada"),
         macroregion = factor(macroregion))
```

Summarize the contents of the dataframe:
```{r}
summary(canada_c19)
```

The earliest date in this dataframe is January 22, 2020, and there are cases at the provincial level by date. The "NAs" are cases attributed to two cruises ("Grand Princess" and "Diamond Princess"), and recovered cases. These can be removed from the dataframe:
```{r}
canada_c19 <- canada_c19 %>%
  drop_na()
```

It is important to note that "Cases" is not the *actual* number of cases (nobody really knows that, although there are estimates of those), but number of *reported* cases. We can see the growth in reported cases by macro-region in the following plot:
```{r}
canada_c19 %>%
  drop_na(macroregion) %>%
  group_by(date, macroregion) %>%
  summarize (Cases = sum(Cases)) %>%
  ggplot(aes(x = date, y = Cases, color = macroregion)) +
  geom_line() +
  theme_bw() +
  ylab("Reported Cases")
```

Quebec has in Canada the largest number of reported cases, followed by Ontario. To calculate the number of new daily cases, we need to lag the Cases variable and substract the previous total:
```{r}
canada_c19 <- canada_c19 %>%
  group_by(Province) %>%
  mutate(new_cases = Cases - lag(Cases, 1)) %>%
  ungroup()
```

The summary of this now shows that the number of new cases ranges from -7 (probably some cases that were retracted) to a maximum of `r max(canada_c19$new_cases)` new daily cases for any province:
```{r}
summary(canada_c19)
```

## Combine the community mobility reports with new COVID-19 cases in Canada

The potential salutary effect of the lockdown (measured in terms of reduced mobility) are not immediate, given the incubation period of the disease. The incubation period has been estimated to be between 2 and 11 days (95% interval) by [Lauer et al. (2020)](https://annals.org/aim/fullarticle/2762808/incubation-period-coronavirus-disease-2019-covid-19-from-publicly-reported). If we wanted to explore the way reduced mobility correlates with number of new cases, we would need to lag the mobility variables by some period. Here, I try two different ways of lagging the mobility variables: a lagged moving average of date-minus-11-days to date-minus-2 days, and a lagged moving *weighted* average of date-minus-11-days to date-minus-2 days (the weights are calculated based on the log-normal distribution reported by Lauer et al.)

First, calculate the weights using the parameters reported by Lauer et al. (2020). For this I will remove the provinces with missing values in their community mobility reports:
```{r}
# Calculate the p-values for the temporal weights
tweights <- data.frame(x = seq(2, 12, 1), p = dlnorm(seq(2, 12, 1), meanlog = 1.621, sdlog = 0.418))

# Sum, to normalize to one
ptotal <- sum(tweights$p)

# Weights are the normalized p-values
tweights <- tweights %>%
  mutate(w = p/ptotal)

tweights <- rbind(data.frame(x = 1, p = 0, w = 0), tweights)
```

Next, calculate the unweighted moving average for each mobility indicator (call this lag *lag11*):
```{r}
# Lagged r_and_r: 11 day moving average
r_and_r_lag11 <- canada_cmr %>% 
  group_by(sub_region_1) %>%
  group_modify(~ {((lag(.x$r_and_r, 12) + 
                      lag(.x$r_and_r, 11) + 
                      lag(.x$r_and_r, 10) + 
                      lag(.x$r_and_r, 9) + 
                      lag(.x$r_and_r, 8) + 
                      lag(.x$r_and_r, 7) + 
                      lag(.x$r_and_r, 6) + 
                      lag(.x$r_and_r, 5) + 
                      lag(.x$r_and_r, 4) + 
                      lag(.x$r_and_r, 3) + 
                      lag(.x$r_and_r, 2))/11) %>%
      enframe()}) %>%
  ungroup() %>% 
  filter(sub_region_1 != "" & 
           sub_region_1 != "Nunavut" &  
           sub_region_1 != "Northwestern Territories" &
           sub_region_1 != "Prince Edward Island" &
           sub_region_1 != "Yukon") %>%
  transmute(sub_region_1 = as.character(sub_region_1), 
            Date = rep(seq(min(canada_cmr$date),
                           max(canada_cmr$date), 
                           by = "days"), 
                       10), 
            Mean_r_and_r_lag11 = value)

# Lagged g_and_p: 11 day moving average
g_and_p_lag11 <- canada_cmr %>% 
  group_by(sub_region_1) %>%
  group_modify(~ {((lag(.x$g_and_p, 12) + 
                      lag(.x$g_and_p, 11) + 
                      lag(.x$g_and_p, 10) + 
                      lag(.x$g_and_p, 9) + 
                      lag(.x$g_and_p, 8) + 
                      lag(.x$g_and_p, 7) + 
                      lag(.x$g_and_p, 6) + 
                      lag(.x$g_and_p, 5) + 
                      lag(.x$g_and_p, 4) + 
                      lag(.x$g_and_p, 3) + 
                      lag(.x$g_and_p, 2))/11) %>%
      enframe()}) %>%
  ungroup() %>% 
  filter(sub_region_1 != "" & 
           sub_region_1 != "Nunavut" & 
           sub_region_1 != "Northwestern Territories" &
           sub_region_1 != "Prince Edward Island" &
           sub_region_1 != "Yukon") %>%
  transmute(sub_region_1 = as.character(sub_region_1), 
            Date = rep(seq(min(canada_cmr$date),
                           max(canada_cmr$date), 
                           by = "days"), 
                       10), 
            Mean_g_and_p_lag11 = value)

# Lagged parks: 11 day moving average
parks_lag11 <- canada_cmr %>% 
  group_by(sub_region_1) %>%
  group_modify(~ {((lag(.x$parks, 12) + 
                      lag(.x$parks, 11) + 
                      lag(.x$parks, 10) + 
                      lag(.x$parks, 9) + 
                      lag(.x$parks, 8) + 
                      lag(.x$parks, 7) + 
                      lag(.x$parks, 6) + 
                      lag(.x$parks, 5) + 
                      lag(.x$parks, 4) + 
                      lag(.x$parks, 3) + 
                      lag(.x$parks, 2))/11) %>%
      enframe()}) %>%
  ungroup() %>% 
  filter(sub_region_1 != "" & 
           sub_region_1 != "Nunavut" & 
           sub_region_1 != "Northwestern Territories" &
           sub_region_1 != "Prince Edward Island" &
           sub_region_1 != "Yukon") %>%
  transmute(sub_region_1 = as.character(sub_region_1), 
            Date = rep(seq(min(canada_cmr$date),
                           max(canada_cmr$date), 
                           by = "days"), 
                       10), 
            Mean_parks_lag11 = value)

# Lagged transit: 11 day moving average
transit_lag11 <- canada_cmr %>% 
  group_by(sub_region_1) %>%
  group_modify(~ {((lag(.x$transit, 12) + 
                      lag(.x$transit, 11) + 
                      lag(.x$transit, 10) + 
                      lag(.x$transit, 9) + 
                      lag(.x$transit, 8) + 
                      lag(.x$transit, 7) + 
                      lag(.x$transit, 6) + 
                      lag(.x$transit, 5) + 
                      lag(.x$transit, 4) + 
                      lag(.x$transit, 3) + 
                      lag(.x$transit, 2))/11) %>%
      enframe()}) %>%
  ungroup() %>% 
  filter(sub_region_1 != "" & 
           sub_region_1 != "Nunavut" & 
           sub_region_1 != "Northwestern Territories" &
           sub_region_1 != "Prince Edward Island" &
           sub_region_1 != "Yukon"  ) %>%
  transmute(sub_region_1 = as.character(sub_region_1), 
            Date = rep(seq(min(canada_cmr$date),
                           max(canada_cmr$date), 
                           by = "days"), 
                       10), 
            Mean_transit_lag11 = value)

# Lagged work: 11 day moving average
work_lag11 <- canada_cmr %>% 
  group_by(sub_region_1) %>%
  group_modify(~ {((lag(.x$work, 12) + 
                      lag(.x$work, 11) + 
                      lag(.x$work, 10) + 
                      lag(.x$work, 9) + 
                      lag(.x$work, 8) + 
                      lag(.x$work, 7) + 
                      lag(.x$work, 6) + 
                      lag(.x$work, 5) + 
                      lag(.x$work, 4) + 
                      lag(.x$work, 3) + 
                      lag(.x$work, 2))/11) %>%
      enframe()}) %>%
  ungroup() %>% 
  filter(sub_region_1 != "" & 
           sub_region_1 != "Nunavut" & 
           sub_region_1 != "Northwestern Territories" &
           sub_region_1 != "Prince Edward Island" &
           sub_region_1 != "Yukon") %>%
  transmute(sub_region_1 = as.character(sub_region_1), 
            Date = rep(seq(min(canada_cmr$date),
                           max(canada_cmr$date), 
                           by = "days"), 
                       10), 
            Mean_work_lag11 = value)

# Lagged residential: 11 day moving average
residential_lag11 <- canada_cmr %>% 
  group_by(sub_region_1) %>%
  group_modify(~ {((lag(.x$residential, 12) + 
                      lag(.x$residential, 11) + 
                      lag(.x$residential, 10) + 
                      lag(.x$residential, 9) + 
                      lag(.x$residential, 8) + 
                      lag(.x$residential, 7) + 
                      lag(.x$residential, 6) + 
                      lag(.x$residential, 5) + 
                      lag(.x$residential, 4) + 
                      lag(.x$residential, 3) + 
                      lag(.x$residential, 2))/11) %>%
      enframe()}) %>%
  ungroup() %>% 
  filter(sub_region_1 != "" & 
           sub_region_1 != "Nunavut" &
           sub_region_1 != "Northwestern Territories" &
           sub_region_1 != "Prince Edward Island" &
           sub_region_1 != "Yukon") %>%
  transmute(sub_region_1 = as.character(sub_region_1), 
            Date = rep(seq(min(canada_cmr$date),
                           max(canada_cmr$date), 
                           by = "days"), 
                       10), 
            Mean_residential_lag11 = value)

```

Next, calculate the weighted moving average for each mobility indicator (call this lag *lag11w*):
```{r}
# Lagged r_and_r: 11 day moving average
r_and_r_lag11w <- canada_cmr %>% 
  group_by(sub_region_1) %>%
  group_modify(~ {((lag(.x$r_and_r, 12) * tweights$w[12] + 
                      lag(.x$r_and_r, 11) * tweights$w[11] + 
                      lag(.x$r_and_r, 10) * tweights$w[10] + 
                      lag(.x$r_and_r, 9) * tweights$w[9] + 
                      lag(.x$r_and_r, 8) * tweights$w[8] + 
                      lag(.x$r_and_r, 7) * tweights$w[7] + 
                      lag(.x$r_and_r, 6) * tweights$w[6] + 
                      lag(.x$r_and_r, 5) * tweights$w[5] + 
                      lag(.x$r_and_r, 4) * tweights$w[4] + 
                      lag(.x$r_and_r, 3) * tweights$w[3] + 
                      lag(.x$r_and_r, 2) * tweights$w[2])/11) %>%
      enframe()}) %>%
  ungroup() %>% 
  filter(sub_region_1 != "" & 
           sub_region_1 != "Nunavut" &  
           sub_region_1 != "Northwestern Territories" & 
           sub_region_1 != "Prince Edward Island" & 
           sub_region_1 != "Yukon") %>%
  transmute(sub_region_1 = as.character(sub_region_1), 
            Date = rep(seq(min(canada_cmr$date),
                           max(canada_cmr$date), 
                           by = "days"), 
                       10), 
            Mean_r_and_r_lag11w = value)

# Lagged g_and_p: 11 day moving average
g_and_p_lag11w <- canada_cmr %>% 
  group_by(sub_region_1) %>%
  group_modify(~ {((lag(.x$g_and_p, 12) * tweights$w[12] + 
                      lag(.x$g_and_p, 11) * tweights$w[11] + 
                      lag(.x$g_and_p, 10) * tweights$w[10] + 
                      lag(.x$g_and_p, 9) * tweights$w[9] + 
                      lag(.x$g_and_p, 8) * tweights$w[8] + 
                      lag(.x$g_and_p, 7) * tweights$w[7] + 
                      lag(.x$g_and_p, 6) * tweights$w[6] + 
                      lag(.x$g_and_p, 5) * tweights$w[5] + 
                      lag(.x$g_and_p, 4) * tweights$w[4] + 
                      lag(.x$g_and_p, 3) * tweights$w[3] + 
                      lag(.x$g_and_p, 2) * tweights$w[2])/11) %>%
      enframe()}) %>%
  ungroup() %>% 
  filter(sub_region_1 != "" & 
           sub_region_1 != "Nunavut" &  
           sub_region_1 != "Northwestern Territories"  & 
           sub_region_1 != "Prince Edward Island"  & 
           sub_region_1 != "Yukon") %>%
  transmute(sub_region_1 = as.character(sub_region_1), 
            Date = rep(seq(min(canada_cmr$date),
                           max(canada_cmr$date), 
                           by = "days"), 
                       10), 
            Mean_g_and_p_lag11w = value)

# Lagged parks: 11 day moving average
parks_lag11w <- canada_cmr %>% 
  group_by(sub_region_1) %>%
  group_modify(~ {((lag(.x$parks, 12) * tweights$w[12] + 
                      lag(.x$parks, 11) * tweights$w[11] + 
                      lag(.x$parks, 10) * tweights$w[10] + 
                      lag(.x$parks, 9) * tweights$w[9] + 
                      lag(.x$parks, 8) * tweights$w[8] + 
                      lag(.x$parks, 7) * tweights$w[7] + 
                      lag(.x$parks, 6) * tweights$w[6] + 
                      lag(.x$parks, 5) * tweights$w[5] + 
                      lag(.x$parks, 4) * tweights$w[4] + 
                      lag(.x$parks, 3) * tweights$w[3] + 
                      lag(.x$parks, 2) * tweights$w[2])/11) %>%
      enframe()}) %>%
  ungroup() %>% 
  filter(sub_region_1 != "" & 
           sub_region_1 != "Nunavut" &  
           sub_region_1 != "Northwestern Territories"  & 
           sub_region_1 != "Prince Edward Island"  & 
           sub_region_1 != "Yukon") %>%
  transmute(sub_region_1 = as.character(sub_region_1), 
            Date = rep(seq(min(canada_cmr$date),
                           max(canada_cmr$date), 
                           by = "days"), 
                       10), 
            Mean_parks_lag11w = value)

# Lagged transit: 11 day moving average
transit_lag11w <- canada_cmr %>% 
  group_by(sub_region_1) %>%
  group_modify(~ {((lag(.x$transit, 12) * tweights$w[12] + 
                      lag(.x$transit, 11) * tweights$w[10] + 
                      lag(.x$transit, 10) * tweights$w[11] + 
                      lag(.x$transit, 9) * tweights$w[9] + 
                      lag(.x$transit, 8) * tweights$w[8] + 
                      lag(.x$transit, 7) * tweights$w[7] + 
                      lag(.x$transit, 6) * tweights$w[6] + 
                      lag(.x$transit, 5) * tweights$w[5] + 
                      lag(.x$transit, 4) * tweights$w[4] + 
                      lag(.x$transit, 3) * tweights$w[3] + 
                      lag(.x$transit, 2) * tweights$w[2])/11) %>%
      enframe()}) %>%
  ungroup() %>% 
  filter(sub_region_1 != "" & 
           sub_region_1 != "Nunavut" &  
           sub_region_1 != "Northwestern Territories"  & 
           sub_region_1 != "Prince Edward Island"  & 
           sub_region_1 != "Yukon") %>%
  transmute(sub_region_1 = as.character(sub_region_1), 
            Date = rep(seq(min(canada_cmr$date),
                           max(canada_cmr$date), 
                           by = "days"), 
                       10), 
            Mean_transit_lag11w = value)

# Lagged work: 11 day moving average
work_lag11w <- canada_cmr %>% 
  group_by(sub_region_1) %>%
  group_modify(~ {((lag(.x$work, 12) * tweights$w[12] + 
                      lag(.x$work, 11) * tweights$w[11] + 
                      lag(.x$work, 10) * tweights$w[10] + 
                      lag(.x$work, 9) * tweights$w[9] + 
                      lag(.x$work, 8) * tweights$w[8] + 
                      lag(.x$work, 7) * tweights$w[7] + 
                      lag(.x$work, 6) * tweights$w[6] + 
                      lag(.x$work, 5) * tweights$w[5] + 
                      lag(.x$work, 4) * tweights$w[4] + 
                      lag(.x$work, 3) * tweights$w[3] + 
                      lag(.x$work, 2) * tweights$w[2])/11) %>%
      enframe()}) %>%
  ungroup() %>% 
  filter(sub_region_1 != "" & 
           sub_region_1 != "Nunavut" &  
           sub_region_1 != "Northwestern Territories"  & 
           sub_region_1 != "Prince Edward Island"  & 
           sub_region_1 != "Yukon") %>%
  transmute(sub_region_1 = as.character(sub_region_1), 
            Date = rep(seq(min(canada_cmr$date),
                           max(canada_cmr$date), 
                           by = "days"), 
                       10), 
            Mean_work_lag11w = value)

# Lagged residential: 11 day moving average
residential_lag11w <- canada_cmr %>% 
  group_by(sub_region_1) %>%
  group_modify(~ {((lag(.x$residential, 12) * tweights$w[12] + 
                      lag(.x$residential, 11) * tweights$w[11] + 
                      lag(.x$residential, 10) * tweights$w[10] + 
                      lag(.x$residential, 9) * tweights$w[9] + 
                      lag(.x$residential, 8) * tweights$w[8] + 
                      lag(.x$residential, 7) * tweights$w[7] + 
                      lag(.x$residential, 6) * tweights$w[6] + 
                      lag(.x$residential, 5) * tweights$w[5] + 
                      lag(.x$residential, 4) * tweights$w[4] + 
                      lag(.x$residential, 3) * tweights$w[3] + 
                      lag(.x$residential, 2) * tweights$w[2])/11) %>%
      enframe()}) %>%
  ungroup() %>% 
  filter(sub_region_1 != "" & 
           sub_region_1 != "Nunavut" &  
           sub_region_1 != "Northwestern Territories"  & 
           sub_region_1 != "Prince Edward Island"  & 
           sub_region_1 != "Yukon") %>%
  transmute(sub_region_1 = as.character(sub_region_1), 
            Date = rep(seq(min(canada_cmr$date),
                           max(canada_cmr$date), 
                           by = "days"), 
                       10), 
            Mean_residential_lag11w = value)
```

Now we can join the lagged mobility indicators to the dataframe with COVID-19 cases:
```{r}
canada_c19 <- canada_c19 %>%
  mutate(Province = as.character(Province)) %>%
  left_join(r_and_r_lag11, by = c("Province" = "sub_region_1", "date" = "Date")) %>%
  left_join(r_and_r_lag11w, by = c("Province" = "sub_region_1", "date" = "Date")) %>%
  left_join(g_and_p_lag11, by = c("Province" = "sub_region_1", "date" = "Date")) %>%
  left_join(g_and_p_lag11w, by = c("Province" = "sub_region_1", "date" = "Date")) %>%
  left_join(parks_lag11, by = c("Province" = "sub_region_1", "date" = "Date")) %>%
  left_join(parks_lag11w, by = c("Province" = "sub_region_1", "date" = "Date")) %>%
  left_join(transit_lag11, by = c("Province" = "sub_region_1", "date" = "Date")) %>%
  left_join(transit_lag11w, by = c("Province" = "sub_region_1", "date" = "Date")) %>%
  left_join(work_lag11, by = c("Province" = "sub_region_1", "date" = "Date")) %>%
  left_join(work_lag11w, by = c("Province" = "sub_region_1", "date" = "Date")) %>%
  left_join(residential_lag11, by = c("Province" = "sub_region_1", "date" = "Date")) %>%
  left_join(residential_lag11w, by = c("Province" = "sub_region_1", "date" = "Date"))
```

A summary of the dataframe after joining the mobility indicators shows that there are numerous "NAs". This is because the coverage of the mobility reports does not coincide with the coverage of JHU regarding COVID-19. The use of lagged moving averages for the mobility indicators means that some days at the beginning of the series of mobility are also lost:
```{r}
summary(canada_c19)
```

Remove all the "NAs" to obtain a clean table with COVID-19 and mobility information:
```{r}
canada_c19 <- canada_c19 %>%
  drop_na()
```

Check that the panel-like structure of the data is balanced - in other words, that there no provinces are missing for any date:
```{r}
table(canada_c19$Province)
```

Nice. So then, we have 51 days of data for 9 provinces.

What was the initial number of cases in each province for this time series?
```{r}
canada_c19 %>% group_by(Province) %>%
  summarize(initial_cases = min(Cases))
```

We can see that the provinces start the time series in more or less the same position, on `r paste(month(min(canada_c19$date), label= TRUE), ". ", day(max(canada_c19$date)), ", ", year(max(canada_c19$date)), sep = "")`.

Create dummy variables for Quebec and Ontario:
```{r}
canada_c19 <- canada_c19 %>%
  mutate(Ontario = ifelse(Province == "Ontario", 1, 0),
         Quebec = ifelse(Province == "Quebec", 1, 0))
```

I will estimate a model with `new_cases` as the dependent variable and residential-based mobility as the independent variable. Why residential-based mobility? All other forms of mobility has collapsed due to restrictions, meaning that for the most part essential travel is centered on the place of residence. It is also a form of mobility that is discretionary, and the only one that has not .

For the analysis, I will expand the coefficient for the mobility variable with the dummy variables for the provinces of Quebec and Ontario, so that:
$$
\beta_m = \beta_{m0} + \beta_{mO} * Ontario + \beta_{mQ} * Quebec
$$

\noindent where $\beta_m$ is the expanded coefficient. These are the interactions for the expanded coefficient:
```{r}
canada_c19 <- canada_c19 %>%
  mutate(mxOntario = Mean_residential_lag11 * Ontario,
         mxQuebec = Mean_residential_lag11 * Quebec)
```

This is the regression model (forcing the regression through the intercept):
```{r}
mod1 <- lm(new_cases ~ 0 + Mean_residential_lag11 +
             mxOntario +
             mxQuebec, 
           data = canada_c19)
summ(mod1)
```

The results indicate that every 1% increase in residential-related mobility from the base level is associated with 1.60 new daily cases. In Ontario, there is an additional effect of 18.11 new daily cases per 1% increase in mobility from the base level, and in Quebec the additional effect is 28.00 new daily cases per 1% increase in mobility from the base line.

This means that the slope for Canada across the board is 1.60; the slope for Ontario is 19.71; and the slope for Quebec is 29.60. Create a prediction dataframe to plot the slopes
```{r}
new.data = expand.grid(Mean_residential_lag11 = seq(-1, 26, by = 0.5), 
                       Region = c("Canada (except Ontario & Quebec)", "Ontario", "Quebec")) %>%
  mutate(mxOntario = ifelse(Region == "Ontario", Mean_residential_lag11, 0),
         mxQuebec =  ifelse(Region == "Quebec", Mean_residential_lag11, 0))

new.data <- data.frame(new.data, new_cases = predict(mod1, newdata = new.data))
```

These are the slopes for the three regions (Canada minus Ontario & Quebec, Ontario, and Quebec)
```{r}
ggplot(data = new.data, 
       aes(x = Mean_residential_lag11, 
           y = new_cases, 
           color = Region,
           linetype = Region)) +
  geom_line(size = 1) +
  theme_light() +
  ylab("Estimated New Daily Cases of COVID-19") +
  xlab("% Change in Residential-based Mobility from Baseline") +
  theme(legend.position = "bottom")
```

This model is rather crude, since it does not consider the panel structure (potential autocorrelation over time and/or space). It is also not straightforward to interpret, because we do not know what the baseline level of mobility was! Nonetheless, it is interesting to see that increases in mobility are associated with more new cases per day, and especially in the two hardest-hit provinces.


