---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Google-COVID-19-Mobility-Reports

<!-- badges: start -->
<!-- badges: end -->

In this notebook I use `R` to retrieve Google's [community mobility reports](https://www.google.com/covid19/mobility/). These reports are being released by Google to support public health experts and the public understand the way mobility has changed with respect to a baseline level for the duration of the COVID-19 pandemic. I was inspired to do this after seeing Theo Lim's GitHub [repository](https://github.com/theochli/google_covid_mobility_map) with Python code to download the reports and visualize the data. This is mostly an example of how to retrieve data, and I will add analysis as I can.

Reports are presented by location with variations; for instance, Canada's community mobility reports are given at the national and provincial levels. For some countries the reports are given only at the national level. For the US the reports are at the level of counties. Each report includes the percent change in visits to places like grocery stores and parks within the designated geographic area. The baseline is the median value, for the corresponding day of the week, during the 5-week period between January, 3rd and February 6th, 2020. Given seasonality in mobility, it would probably have been better to use a baseline of the corresponding day of the week for a moving window of 5 or whatever weeks, but of the previous year. I wonder if those data from previous years even exist.

The categories of places are as follows:

- Grocery & pharmacy: Mobility trends for places like grocery markets, food warehouses, farmers markets, specialty food shops, drug stores, and pharmacies.

- Parks: Mobility trends for places like local parks, national parks, public beaches, marinas, dog parks, plazas, and public gardens.

- Transit stations: Mobility trends for places like public transport hubs such as subway, bus, and train stations.

- Retail & recreation: Mobility trends for places like restaurants, cafes, shopping centers, theme parks, museums, libraries, and movie theaters.

- Residential: Mobility trends for places of residence.

- Workplaces: Mobility trends for places of work.

More information about the data can be found [here](https://www.google.com/covid19/mobility/data_documentation.html?hl=en).

## Preliminaries

Load the packages needed to read and work with the data:
```{r message=FALSE}
library(lubridate)
library (RCurl)
library(tidyverse)
```

Retrieve data:
```{r}
download <- getURL("https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv")
gmobility <- read.csv (text = download)
```

Save as RData. I created a folder called `data` to keep a local copy of the mobility report:
```{r}
save(gmobility, file = "data/gmobility.RData")
```

What are the contents of the report?
```{r}
head(gmobility)
```

The unit of the report is a combination of region and date, and the columns include the percent change from baseline for each category of place.

Check the range of dates covered:
```{r}
summary(gmobility$date)
```

The earliest date reported for any geographical location February 15, 2020. However, the `date` is stored as a factor which is not ideal work with times; I will change it to date:
```{r}
gmobility <- gmobility %>%
  mutate(date = lubridate::ymd(date))
```

Summarize the dates again:
```{r}
summary(gmobility$date)
```

That is better. Now it is clear that the earliest date is Feb. 15, 2020 (the minimum value) and the latest (for the time being) is `r paste(month(max(gmobility$date), label= TRUE), ". ", day(max(gmobility$date)), ", ", year(max(gmobility$date)), sep = "")`.

With this more or less the community reports are in workable shape.

## How has mobility in Canada changed according to Google?

To begin an exploration of mobility in Canada (according to Google), the first step is to filter the dataframe to include only Canadian data (I'm calling this "cmr" for "community mobility report"):
```{r}
canada_cmr <- gmobility %>%
  filter(country_region == "Canada") %>%
  droplevels()
```

We can quickly check the contents of this Canadian dataframe:
```{r}
summary(canada_cmr)
```

Now we have an All-Canadian dataset with the highest geographical level of resolution the province. At this point, I am going to rename the annoyingly long names of the columns to something more practical (after all, I know that all these variables are "percent change from baseline"):
```{r}
canada_cmr <- canada_cmr %>%
  rename(r_and_r = retail_and_recreation_percent_change_from_baseline, 
         g_and_p = grocery_and_pharmacy_percent_change_from_baseline, 
         parks = parks_percent_change_from_baseline,
         transit = transit_stations_percent_change_from_baseline, 
         work = workplaces_percent_change_from_baseline,
         residential = residential_percent_change_from_baseline)
```

Another thing that we notice is that there are quite a few "NAs" in the table. These could be data that have been censored by Google to comply with their [Privacy Policy](https://www.google.com/covid19/mobility/index.html?hl=en). Maybe. But that seems like a little bit of a stretch, given that the aggregation is at the provincial level. Or it could be that in fact there are relatively few people in some provinces at certain times who allow Google to track them. Or could the "NAs" be that on certain dates there were absolutely no trips to those places? What would be the percentage change from the baseline to zero?

We can explore this question by looking at where the "NAs" are:
```{r}
canada_cmr %>% 
  filter(is.na(r_and_r)) %>%
  select(sub_region_1) %>%
  table()
```

Ah. So the "NAs" are concentrated in the Northwest Territories, Nunavut, and Yukon, all sparsely populated regions; likely the "NAs" then are a result of censoring due to small numbers. Notice that there is also an unlabeled "region". This are in fact the national-level statistics.

We can verify that this is the case too with the other categories of places, like groceries and pharmacies:
```{r}
canada_cmr %>% 
  filter(is.na(g_and_p)) %>%
  select(sub_region_1) %>%
  table()
```

Parks:
```{r}
canada_cmr %>% 
  filter(is.na(parks)) %>%
  select(sub_region_1) %>%
  table()
```

In addition to the three regions identified before, Prince Edward Island also has "NAs" in the category of parks.

Transit:
```{r}
canada_cmr %>% 
  filter(is.na(transit)) %>%
  select(sub_region_1) %>%
  table()
```

Work:
```{r}
canada_cmr %>% 
  filter(is.na(work)) %>%
  select(sub_region_1) %>%
  table()
```

Residential:
```{r}
canada_cmr %>% 
  filter(is.na(work)) %>%
  select(sub_region_1) %>%
  table()
```

Now, there are different ways in which these data could be analyzed (obviously), but I will begin with data at the national level. I could remove the provinces/territories with "NAs", or I could include them and ignore the records that are missing. I'll do the latter.

Create a plot of the changes in mobility with respect to retail and recreation, taking the daily average of all provinces (taking care of removing the "NAs" from the calculation of the mean, so using na.rm = TRUE):
```{r}
canada_cmr %>%
  group_by(date) %>%
  summarize (r_and_r = mean(r_and_r, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = r_and_r)) +
  geom_line() +
  theme_bw() +
  ylab("% change wrt baseline: retail and recreation")
```

We see that, at the national level, the mobility hovered close to the baseline (percent change ~ 0) until March 13, which is when social distancing measures started to be phased in across the country. Mobility with respect to retail and recreation after that date collapsed to about -50% change with respect to the baseline, and hovered there (with additional drops in the second week of April).

But did this change similarly across the country? Repeat the plot, but by macro-region. I will add the macro-regions to the dataframe:
```{r}
canada_cmr <- canada_cmr %>%
  mutate(macroregion = case_when(sub_region_1 == "Ontario" ~ "Ontario",
                                 sub_region_1 == "Quebec" ~ "Quebec",
                                 sub_region_1 == "British Columbia" ~ "British Columbia",
                                 sub_region_1 == "Manitoba" | 
                                   sub_region_1 == "Saskatchewan" | 
                                   sub_region_1 == "Alberta"  ~ "Prairies",
                                 sub_region_1 == "New Brunswick" | 
                                   sub_region_1 == "Newfoundland and Labrador" | 
                                   sub_region_1 == "Nova Scotia" | 
                                   sub_region_1 == "Prince Edward Island"  ~ "Atlantic",
                                 sub_region_1 == "Northwest Territories" | 
                                   sub_region_1 == "Nunavut" | 
                                   sub_region_1 == "Yukon" ~ "Northern Canada"),
         macroregion = factor(macroregion))
```

```{r}
canada_cmr %>%
  drop_na(macroregion) %>%
  group_by(date, macroregion) %>%
  summarize (r_and_r = mean(r_and_r, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = r_and_r, color = macroregion)) +
  geom_line() +
  theme_bw() +
  ylab("% change wrt baseline: retail and recreation")
```

Here, we see that Quebec has had the largest changes in retail-and-recreation-related mobility of all regions, whereas Northern Canada has had the least (albeit still quite considerable)(but we don't know what was the baseline; if the base line was low mobility, then there was maybe not much that could be cut).







